---
title: 'PH142 Data Project'
author: "Colin FitzGerald, Natalie Miramontes, Quinn Dahl, Amanda Tanaka, Celina Alnakib"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(broom)
library(dplyr)
```

1. [2 marks] What is the problem your are addressing with these data? State the question you are trying
to answer and let us know what type of question this is in terms of the PPDAC framework.

The problem we are addressing with this data is the estimation of obesity levels based on eating habits and physical condition. 

We are trying to answer the question of whether there are specific factors within eating habits and physical conditions that leads to a higher levels of obesity. Some of these factors include frequent consumption of high caloric food, number of main meals, consumption of water daily, physical activity frequency, and transportation used. This question is a causative/etiologic question:

Are there potential associations or trends within the eating habits and physical conditions of individuals of Mexico, Peru and Columbia that may have lead to higher levels of obesity in the respective countries?

\newpage

2. [2 marks] What is the target population for your project? Why was this target chosen i.e., what was
your rationale for wanting to answer this question in this specific population?

The target population for this project are individuals from Mexico, Peru, and Columbia between the ages of 14 and 61 with diverse eating habits and physical conditions. 

The rationale for wanting to answer this question in this specific population is that  we want to study potential community or cultural factors that may have led to high obesity levels and be able to make a generalization of risk factors for these specific countries.

There are people with all different factors and exposures that can help us identify key factors that lead to an increase in obesity levels. Specifically, these countries were chosen as they have high rates of obesity and are currently showing no steps at lowering them. This may provide insight on the factors that increase obesity levels and may show us areas that can be focused on to lower these levels. 

\newpage

3. [2 marks] What is the sampling frame used to collect the data you are using? Describe why you think
this sampling strategy is appropriate for your question. To what group(s) would you feel comfortable
generalizing the findings of your study and why.

This data comes from an estimation of obesity levels in people from the countries of Mexico, Peru and Colombia, with ages between 14 and 61 and diverse eating habits and physical condition. 

Data was collected using a web platform with a survey where anonymous users answered each question, then the information was processed obtaining 17 attributes and 2111 records. The attributes that were collected were related to eating habits, physical conditions, gender, age, height and weight. 
This sampling strategy is appropriate for our question given we are interested in looking at and getting more information on potential community or culture driven factors that lead to a higher levels of obesity in each country. 

We would feel comfortable generalizing the findings of our study to Mexico, Peru, and Colombia’s neighboring countries in Central and South America given that we may find potential trends and associations in the eating and physical habits across the three countries.

\newpage

4. [2 marks] Write a brief description (1-4 sentences) of the source and contents of your dataset. Provide
a URL to the original data source if applicable. If not (e.g., the data came from your internship),
provide 1-2 sentences saying where the data came from. If you completed a web form to access the
data and selected a subset, describe these steps (including any options you selected) and the date you
accessed the data.

https://archive.ics.uci.edu/ml/datasets/Estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition+. 

The data came from the UC Irvine Machine Learning Repository and contains information such as Frequent consumption of high caloric food (FAVC), Frequency of consumption of vegetables (FCVC), Physical activity frequency (FAF), Time using technology devices (TUE), Transportation used (MTRANS), and other variables as such. We didn’t have to complete a web form to access the data and did not need to subset it either. There were links to download both the Data Folder and the Data Set Description at the top of the website. We accessed the data on Monday February 21, 2022. 

\newpage

5. [1 mark] Write code below to import your data into R. Assign your dataset to an object.
```{r, message=FALSE}
obesity_data <- read_csv("ObesityDataSet_raw_and_data_sinthetic.csv")
```

\newpage

6. [3 marks] Use code in R to answer the following questions:
i) What are the dimensions of the dataset?

```{r}
dim(obesity_data)
```

The dimensions of the dataset are 2111 rows by 17 columns. 


ii) Provide a list of variable names.

```{r}
names(obesity_data)
```


iii) Print the first six rows of the dataset.

```{r}
head(obesity_data)
```

\newpage


7. [4 marks] Use the data to demonstrate a statistical concept from Part I of the course. Describe the
concept that you are demonstrating and interpret the findings. This should be a combination of code
and written explanation.

The overarching statistical concept that our group wanted to demostrated in the linear association between two quantitative variables from our dataset. The two variables we did this with were height of participants (in meters) and weight of participants (in kilograms). The method we used to do so was performing linear regression of the dependent variable in this case of weight on the explanatory variable of height. We first made a scatter plot through ggplot2 with a best fit regression line include from the first code chunk below to visualize the data and get a sense of what the regression coefficients should look like. We then performed linear regression shown in the second code chunk below and received a readout for the regression coefficients as -134.6 for the intercept coefficient and 130.0 for the coefficient of height on weight. In context, the intercept coefficient indicates that a person with a height of 0.0 meters should weigh -134.6 kg. Obviously this doesn't make sense in context because people cannot have negative weights and this readout comes from the narrowness of our sampling data. Only people between roughly 1.4 meters and 2.0 meters in height are included in this dataframe, and because this data is looking at primarily obese individuals, the data most likely has a somewhat unrepresentative height to weight ratio when performing regression as compared to the general population. In context, the readout for the height coefficient suggests that a 1 meter increase in the height of a participant will mean that there will be a 130 kg increase in the weight of the individual. This coefficient makes more sense in context as there tends to be a direct relationship between height and weight in people, but the magnitude of this coefficient is likely biased to the selection of the sampling frame as well.

```{r}
obesity_hw_scatter = ggplot(data = obesity_data, aes(x = Height, y = Weight)) + 
  geom_point() +
  geom_smooth(method = lm) + 
  labs(title = 'Obesity Data Heights and Weights Plotted')
obesity_hw_scatter
```

```{r}
obesity_hw_regression <- lm(Weight ~ Height, data = obesity_data)
obesity_hw_regression
```


\newpage 

## Part 2

2. [2 marks] Describe a quantity you will estimate as an outcome in your problem using probability
notation. Are you planning to calculate marginal probabilities? Conditional probabilities?

```{r}
library(janitor)
obesity_data$CAEC <- factor(obesity_data$CAEC, levels=c('no', 'Sometimes', 'Frequently', 'Always'))
tabyl(obesity_data, Gender, CAEC) %>% adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages("row") %>% adorn_pct_formatting() %>% adorn_ns() %>% knitr::kable()
```

We would like to calculate the conditional probabilities of eating habits based on gender. We are curious to see how eating habits between meals vary by gender. The above table creates those conditional probabilities for us. The quantities we will estimate as an outcome, are categorical variables, (no, sometimes, frequently, always), based on gender. Below we report those calculated probabilities. These are estimations for this data set only. 

$P$(does not eat between meals | Female) $= 15/1043 = 1.4$%

$P$(sometimes eats between meals | Female) $= 844/1043 = 80.9$%

$P$(frequently eats between meals | Female) $= 161/1043 = 15.4$%

$P$(always eats between meals | Female) $= 23/1043 = 2.2$%

$P$(does not eat between meals | Male) $= 36/1068 = 3.4$%

$P$(sometimes eats between meals | Male) $= 921/1068 = 86.2$%

$P$(frequently eats between meals | Male) $= 81/1068 = 7.6$%

$P$(always eats between meals | Male) $= 30/1068 = 2.8$%

The above probabilities are the probability of each eating habit conditioned on gender. 

\newpage


3. [3 marks] Describe the type of theoretical distribution that is relevant for your data.

i) What type of variable(s) are you investigating (continuous, categorical, ordinal, etc)?

There are 17 attributes in our data set and all of the variables were categorical except three: Weight, Height, and Age. Those three variables were numerical and continuous since we can assume an infinite number of real values within a given interval.
It is worth noting that our data set also includes 5 binary attributes since each observation either falls into YES or NO categories or in the case of Gender ( MALE or FEMALE), there are a fixed number of observations, each observation is independent, and the probability of each observation is 50%: 

Gender
Frequent consumption of high caloric food (FAVC)
Calories consumption monitoring (SCC)
Smoker
Familial Overweight Connection


Some other variables of interest are Time using technology devices (TUE), Number of main meals (NCP), Consumption of food between meals (CAEC), which are all categorical and ordinal variables in our data set. This is the case since their values are defined by an order relation between the different categories ie. sometimes, frequently, and always.

ii) What theoretical distribution that we have talked about would potentially be appropriate to use with these data (Normal, Binomial, Poisson. . . )

If we were to investigate one of the continuous variables such as height or age, then a normal distribution would be appropriate. 
However, if we were to investigate one of the binary variables then a binomial distribution would be more appropriate, which could be modeled from a Bernoulli variable derived from setting 'yes' and 'no' to 0 and 1 in the data set. 

The main distributions for categorical data analysis are the binomial and multinomial distributions. Probabilities associated with different combinations of events can be determined based on the distributions.

iii) Why is this an appropriate model for the data you are studying?

Let’s focus on investigating height for the purpose of this question. In order to check whether or not the normal distribution is appropriate, we plotted a QQ plot. 

```{r}
ggplot(obesity_data, aes(sample = Height)) + stat_qq() + stat_qq_line() +
theme_minimal(base_size = 15) + 
  labs(title = "QQ Plot of Height from Obesity Data Set") + 
  xlab("Theoretical") + ylab("Sample")
```

When we plot the theoretical quantiles on the Q-Q plot, we can see that the points are most closely attached from about a height of 1.6 to 1.8 meters so a normal distribution is appropriate for those values; however, as we go higher and lower we start to see a little bit of a skew so a normal distribution would no longer be appropriate. 

Furthermore, after stratifying height by gender, we plotted their histograms and we can see that they follow a relatively normal distribution for both genders. 

```{r}
ggplot(obesity_data,aes(x = Height, colour=Gender)) + 
  geom_histogram(aes(y = ..count.., fill=Gender), alpha=0.5,position="identity") + 
  ggtitle("Histogram of Height by Gender from Data Set")
```

\newpage


4. [4 marks] Use the data you have to demonstrate a statistical concept from Part II of the course.
Describe the concept that you are demonstrating and interpret the findings. This may include code in
R, a visual of some kind and text interpretation

```{r}
mean_height <- mean(obesity_data$Height)
std_height <- sd(obesity_data$Height)

men <-obesity_data %>% filter(Gender == 'Male')
male_mean_height <- mean(men$Height)
std_male_height <- sd(men$Height)

women <- obesity_data %>% filter(Gender == 'Female')
female_mean_height <- mean(women$Height)
std_female_height <- sd(women$Height)
```

```{r}
CI_lower_all_data <- mean_height - (1.96 * (std_height/sqrt(nrow(obesity_data))))
CI_upper_all_data <- mean_height + (1.96 * (std_height/sqrt(nrow(obesity_data))))

CI_lower_male_data <- male_mean_height - (1.96 * (std_male_height/sqrt(nrow(men))))
CI_upper_male_data <- male_mean_height + (1.96 * (std_male_height/sqrt(nrow(men))))

CI_lower_female_data <- female_mean_height - (1.96 * (std_female_height/sqrt(nrow(women))))
CI_upper_female_data <- female_mean_height + (1.96 * (std_female_height/sqrt(nrow(women))))
```

```{r}
CI_all_data <- c(CI_lower_all_data, CI_upper_all_data)
CI_all_data
```

```{r}
CI_male_data <- c(CI_lower_male_data, CI_upper_male_data)
CI_male_data
```

```{r}
CI_female_data <- c(CI_lower_female_data, CI_upper_female_data)
CI_female_data
```

```{r}
library(ggplot2)
library(MASS)       # for fitdistr(...)
get.params <- function(z) with(fitdistr(z,"normal"),estimate[1:2])
df <- aggregate(Height~Gender, obesity_data, get.params)
df <- data.frame(Gender=df[,1],df[,2])
x  <- with(obesity_data, seq(min(obesity_data$Height),max(obesity_data$Height),len=100))
gg <- data.frame(Height=rep(x,nrow(df)),df)
gg$y <- with(gg,dnorm(x,mean,sd))
gg$y <- gg$y * aggregate(Height~Gender, obesity_data,length)$Height * diff(range(obesity_data$Height))/30

ggplot(obesity_data,aes(x = Height, colour=Gender)) + 
  geom_histogram(aes(y = ..count.., fill=Gender), alpha=0.5,position="identity") +
  geom_line(data=gg, aes(y=y)) + 
  geom_vline(xintercept = CI_lower_female_data) + 
  geom_vline(xintercept = CI_upper_female_data) + 
  geom_vline(xintercept = CI_lower_male_data) + geom_vline(xintercept = CI_upper_male_data)
```

The above is a visualization of our confidence intervals, for the stratification by gender. 

\newpage

The concept that we are demonstrating is finding a confidence interval. A confidence interval displays the probability that a parameter will fall between a pair of values around the mean. Specifically within our data, we found the confidence interval for our entire dataset and stratified our data by gender (male and female). Our confidence intervals specifically work to capture the mean height of obese males, females, and both in Mexico, Peru and Columbia. Based on our results, we found that the interval 1.697697 - 1.705658 (meters) has a 95% success rate in capturing within that interval the mean height µ of all obese males and females in Mexico, Peru and Columbia. After we stratified our data by gender we found that the interval 1.754362 - 1.763019 (meters) has a 95% success rate in capturing within that interval the mean height µ of obese males in Mexico, Peru and Columbia. We also found that the interval 1.638776 - 1.647820 (meters) has a 95% success rate in capturing within that interval the mean height µ of obese females in Mexico, Peru and Columbia. By conducting these confidence intervals, we found there to be a statistically significant difference in the true mean height of all 3 intervals due to there not being any overlap in confidence intervals. 

\newpage

## Part 3

2. [2 points] Identify a statistical test to apply to your data (must be a concept we covered in part III of the course). In plain language, write the question you are trying to answer. 


The statistical test we plan to apply to our data is a chi-square test for independence. The question we are trying to answer is if there is an association between obesity level and gender within individuals of our study. This is to analyze if gender is a risk factor for obesity. We are using 2 categories of gender (male and female) and 7 obesity levels (insufficient weight, normal weight, obesity type 1, obesity type 2, obesity type 3, overweight level 1, overweight level 2) to see if there is an association between level and gender. 

\newpage


3. [2 points] What assumptions are required by the method you chose in question 2? Show how you assessed whether these assumptions are met by your dataset. 


An assumption that is required by the chi-squared test is to have performed a simple random sample to obtain the data. This data was collected by an online survey that may have not necessarily been random as we are prone to bias of response and nonresponse. Therefore this is not a complete random sample, but for this data project, we will continue with its use. Another assumption is that the variables are categorical and we made sure to choose 2 categorical variables. With the obesity data, we made sure not to use weight as the determining factor because that would be a continuous variable, but the obesity levels are categorical. The last assumption that our data met to proceed with this test is that the expected counts are at least 5 which we see to be the case in our expected table that we calculated. 

\newpage


4. [2 points] Explain why this test is appropriate for the data you have and the question you are trying to answer. Use at least one visualization technique and include both the output and the R code that generated it. 


A chi-square test for independence is appropriate for our data because we are using 2 categorical variables that are able to be counted easily. We are looking to see if there is a relationship between the two and this test would allow us to determine if there was one. We see with our data in the 2 way table that it is in good format for counts with a chi-squared test and visualization with a dodge bar chart. 

```{r}
ggplot(obesity_data,aes(x = NObeyesdad, colour = Gender)) + 
  geom_bar(aes(y = ..count.., fill=Gender), alpha=0.5,position = "dodge") + 
  ggtitle("Barplot of Obesity Levels by Gender from Data Set") + theme(axis.text.x = element_text(angle = 45, hjust=1))
```
```{r}
ggplot(obesity_data,aes(x = NObeyesdad , colour = Gender))+ theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  geom_bar(aes(y = ..count.., fill=Gender, stat="identity")) + facet_grid(.~FAVC) 

```

```{r}
library(janitor)
observed_table <- matrix(c(173, 141, 156,2,323,145,103, 99,146,195,295,1,145,187), 
                         nrow = 2, ncol = 7, byrow = T)
rownames(observed_table) <- c('Female', 'Male')
colnames(observed_table) <- c('Insufficient Weight', 
                              'Normal Weight', 
                              'Obesity Type I', 
                              'Obesity Type II',
                              'Obesity Type III',
                              'Overweight Level I',
                              'Overweight Level II')
ObservedVal_Table <- knitr::kable(observed_table, caption = "Observed Data")
ObservedVal_Table
ExpectedVal_Table <- knitr::kable(chisq.test(observed_table)$expected, 
                                  caption = "Expected Data")
ExpectedVal_Table
```
\newpage

5. [2 points] Clearly state the null and alternative hypotheses for this test. 

$H_{o}$: Gender and obesity level are independent. 
$H_{a}$: Gender and obesity level are not independent. 

\newpage


6. [2 points] Include the R code you used to generate your results. Annotate your code to help us follow your reasoning. 

```{r}
#perform chi-squared test in order to generate test statistic
chisq.test(observed_table)
#use the chi-squared test statistic in order to 
#calculate the p-value with the appropriate 
#degrees of freedom (rows - 1)(columns-1) = 6
pchisq(q=657.75, df=6, lower.tail = FALSE)
```
\newpage

7. [4 points] Present your results in a clear summary. This should include both a text summary and a table or figure with appropriate labeling. 

Assuming no association between gender and obesity, there is approximately a 0% chance of the chi-square value we calculated or a larger one. This probability is small enough that there is evidence in favor of the alternative hypothesis that there is a relationship between gender and obesity.

Below we created a visualization of the distribution of each weight class by each gender. This table was created using the observed values in the survey.  

```{r}
library(DescTools)
#gender_obesity_table <- tabyl(obesity_data, Gender, NObeyesdad) %>% knitr::kable()
#ExpectedVal_Table
#mosaicplot(gender_obesity_t, dir = c("h", "v"))
#mosaicplot(obesity_data)
PlotMosaic(observed_table, main = "NObeyesdad ~ Gender", 
           off = 0.05, horiz = FALSE, xlab = "NObeyesdad", y = "Gender")
```

\newpage 


8. [4 points] Interpret your findings. Include a statement about the evidence, your conclusions, and the generalizability of your findings. Our analysis and conclusions depend on the quality of our study design and the methods of data collection. Any missteps or oversights during the data collection process could potentially change the outcome of what we are trying to find. Consider the methods used to collect the data you analyzed. Was there any potential issue in how the participants were selected/recruited, retained, or assessed that may have impacted the outcome of your analysis/visualization? Were there any potential biases that you might be concerned about? Were there factors that were not measured or considered that you think could be important to the interpretation of these data? 



According to the tables we generated for observed and expected counts of gender-obesity status pairs, there were large discrepancies between the observed and expected counts for many data pairs which is most likely the reason we observed such an extreme chi-square test statistic and p-value. Based on the chi-square test results, the data suggests that there is dependence between the variables of gender and obesity status. However, we are not particularly optimistic of the generalizability of these findings due to the methods of data collection in order to generate our data. The data was generated from an anonymous online survey which could lead to two major data collection issues. The first is related to the bias of those who chose to participate. If the survey was targeted at viewing associations of obesity indicators to people and their obesity status, it may have been more applicable for individuals who fall under the obesity spectrum to participate. Additionally, even though the survey was anonymous, people of certain groups may be more or less likely to respond honestly to the survey given any views on their health status.



\newpage


9. [1 point] Create a statement of contribution. For example, the American Journal of Epidemiology provides the following instructions to authors: “Authorship credit should be based on criteria developed by the International Committee for Medical Journal Editors (ICMJE): 1) substantial contributions to conception and design, or acquisition of data, or analysis and interpretation of data; 2) drafting the article or reviewing it and, if appropriate, revising it critically for important intellectual content; 3) final approval of the version to be published. Authors should meet all conditions. In addition, each author must certify that they have participated sufficiently in the work to believe in its overall validity and to take public responsibility for appropriate portions of its content. Author names should be listed in ScholarOne and author contributions should be detailed in the cover letter (e.g., “Author A designed the study and directed its implementation, including quality assurance and control. Author B helped supervise the field activities and designed the study’s analytic strategy. Author C helped conduct the literature review and prepare the Methods and the Discussion sections of the text.”).”


Every group member contributed relatively equally to the overarching project given we had joint responsibilities for a lot of the questions.
Member QD had joint responsibility for question 7 in Part I of the project, question 4 in Part II, and questions  6-8 in Part III.

Member CA had joint responsibility for questions 3 and 4 in Part I of the project, question 3 in Part II of the project, and questions 2, 3, 4 and 5 in Part III of the project.
Member AT had joint responsibility for questions 1 and 2 in Part I of the project, question 4 in Part II of the project, and questions 2, 3, 4 and 5 in Part III of the project.

Member NM had joint responsibility for questions 3 and 4 in Part I of the project, question 3 in Part II, and questions 6-8 in Part III.

Member CF had joint responsibility for questions 5 and 6 in Part I of the project, question 2 and 3 in part II of the project, and questions 6-8 in Part III. 
 




