---
title: 'PH142 Data Project'
author: "Colin FitzGerald, Natalie Miramontes, Quinn Dahl, Amanda Tanaka, Celina Alnakib"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(broom)
library(dplyr)
```

1. [2 marks] What is the problem your are addressing with these data? State the question you are trying
to answer and let us know what type of question this is in terms of the PPDAC framework.

The problem we are addressing with this data is the estimation of obesity levels based on eating habits and physical condition. We are trying answer the question of whether there are specific factors within eating habits and physical conditions that leads to a higher levels of obesity. Some of these factors include frequent consumption of high caloric food, number of main meals, consumption of water daily, physical activity frequency, and transportation used. This question is a causative/etiologic question. 


2. [2 marks] What is the target population for your project? Why was this target chosen i.e., what was
your rationale for wanting to answer this question in this specific population?

The target population for this project are individuals from Mexico, Peru, and Columbia between the ages of 14 and 61 with diverse eating habits and physical conditions. The rationale for wanting to answer this question in this specific population is that we have a wide variety of the population in which making a generalization would be more accurate. There are people with all different factors and exposures that can help us identify key factors that lead to an increase in obesity levels. Specifically, these countries were chosen as they have high rates of obesity and are currently showing no steps at lowering them. 

3. [2 marks] What is the sampling frame used to collect the data you are using? Describe why you think
this sampling strategy is appropriate for your question. To what group(s) would you feel comfortable
generalizing the findings of your study and why.

# answer here


4. [2 marks] Write a brief description (1-4 sentences) of the source and contents of your dataset. Provide
a URL to the original data source if applicable. If not (e.g., the data came from your internship),
provide 1-2 sentences saying where the data came from. If you completed a web form to access the
data and selected a subset, describe these steps (including any options you selected) and the date you
accessed the data.

# answer here 


5. [1 mark] Write code below to import your data into R. Assign your dataset to an object.
```{r, message=FALSE}
obesity_data <- read_csv("ObesityDataSet_raw_and_data_sinthetic.csv")
```

6. [3 marks] Use code in R to answer the following questions:
i) What are the dimensions of the dataset?

```{r}
dim(obesity_data)
```

The dimensions of the dataset are 2111 rows by 17 columns. 


ii) Provide a list of variable names.

```{r}
spec(obesity_data)
```


iii) Print the first six rows of the dataset.

```{r}
head(obesity_data)
```


7. [4 marks] Use the data to demonstrate a statistical concept from Part I of the course. Describe the
concept that you are demonstrating and interpret the findings. This should be a combination of code
and written explanation.

# answer here 



